
---
title: "Two-Way-Fixed-Effect-Model"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Two-Way-Fixed-Effect-Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(xtreg2way)
```

The `xtreg2way` package is an algorithm to efficiently estimate a two-way fixed effects model.  This algorithm is adapted from the [Matlab code](https://soma.people.stanford.edu/research) written for [this paper](https://soma.people.stanford.edu/sites/g/files/sbiybj4171/f/jem-2014-0008.pdf).  

To begin, we will generate sample data. It's likely you already have data if you wish to use this package; Skip to the bottom to see how this is used.  First, we decide how many groups and time effects we want.  These are the two effects in the "two-way fixed effects" model.
```{r}
numgroups <- 1000
T <- 200
```

Next, we will assign each observation to a group and time effect.  These are stored in `hhid` and `tid`.  The variable `w` is a weight for each observation.
```{r}
observations <- numgroups * T
e <- 1:observations
## Create groups and weights
hhid <- floor((e - 1) / T + 1)
tid <-  e - (hhid - 1) * T
w <- pracma::rand(n = numgroups, m = 1)
w <- w[hhid]
```

Now each observation has a group, time factor, and weight.  Next, we will generate two random `x` values as our independant variables, based on these effects.  
```{r}
#Randomly create effects for groups
heffect <- pracma::randn(n = numgroups, m = 1)
teffect <- pracma::randn(n = T, m = 1)
#Generate independant variables 
x1 <- pracma::randn(n = observations, m = 1) + 
  0.5 * heffect[hhid] + 0.25 * teffect[tid]
x2 <- pracma::randn(n = observations, m = 1) - 
  0.25 * heffect[hhid] + 0.5 * teffect[tid]
```

Finally, we will generate a random error term, and then create our `y` dependant variable from the variables created.
```{r}
#Generate Random Error
autoc <- pracma::rand(n = numgroups, m = 1)
initialv <- pracma::randn(n = numgroups, m = 1)
u <- pracma::randn(n = observations, m = 1)
for (o in 1:observations) {
  if (tid[o] > 1){
    u_1 <- u[o-1]
  } else {
    u_1 <- initialv[hhid[o]]
  }
  u[o] <- autoc[hhid[o]] * u_1 + u[o]
}
# Generate dependant variable
y <- 1 + x1 - x2 + heffect[hhid] + teffect[tid] + u
```

The function `xtreg2way` has two uses.  The first time you use it, you must specify, `y`,`X`,`hhid`, and `tid`.  The weights vector `w` is assumed to be all 1's if omitted.  
```{r}
#XTREG2WAY
output <- xtreg2way(y, data.frame(x1,x2), hhid, tid, w, noise="1")
```
Using the `noise="1"` parameter, we can see the regression coefficients for `x1` and `x2` printed here, as well as standard errors, t-stat values, and p-values.  Additionally, the `output` variable has several objects.  \

`betaHat`  is a vector of the regression coefficients.
`aVarHat` is the asymptotic variance of the estimator
`y` and `X` are the input variables projected onto the two-way dummy matrices.
`struc` contains several matrices necessary to perform this regression again on different independant variables, for the same set of observations.


The second way to use `xtreg2way` is to use the `struc` object from the first run, but now on a different set of columns.  Below, we see the same process but only using `x1` now.
```{r}
#XTREG2WAY second time
output2 <- xtreg2way(y, x1, struc=output$struc)
```

Providing `struc` saves computation time, making the algorithm much faster.  The output variable `output2` now only contains `betaHat` and `aVarHat`.  